---
title: "R Notebook"
output: html_notebook
---
1. Illustration of mean and standard deviation.
(a) Compute the mean and the standard deviation for 1, 1.5, 2, 2.5, 3.
(b) Compute the mean and the standard deviation for 1, 1.5, 2, 2.5, 30.
(c) Comment on the differences.

```{r Task 1}
# a) ====
x1 <-c(1, 1.5, 2, 2.5, 3)
x1_mean<- mean(x1)
x1_sd<-sd(x1)
#b)====
x2<-c(1, 1.5, 2, 2.5, 30)
x2_mean<-mean(x2)
x2_sd<-sd(x2)
#c)====
x1_mean
x2_mean
x1_sd
x2_sd
#the mean is heavily influenced by  extreme values in the dataset and is not a robust against spread

```
2. Comparing normality for two genes. Consider the gene expression val-
ues in row 790 and 66 of the Golub et al. (1999) data.
(a) Produce a boxplot for the expression values of the ALL patients
and comment on the differences. Are there outliers?
For line 790 there are 3 outliers, for line 66 there are none
(b) Produce a QQ-plot and formulate a hypothesis about the normal-
ity of the genes.
66 is normally distributed, but 790 is not
(c) Compute the mean and the median for the expression values of
the ALL patients and compare these. Do this for both genes.
```{r Task 2}
library(ALL)
data(golub, package = "multtest")
i<-c(66,790)
#a) ====
data_snippet_790<-golub[790,]
data_snippet_66<- golub[66,]
gol.fac <- factor(golub.cl,levels=0:1, labels= c("ALL","AML"))
boxplot(golub[790,],gol.fac=="ALL")
boxplot(golub[66,],gol.fac=="ALL")
boxplot(golub[790,]~gol.fac)
boxplot(golub[66,]~gol.fac)
#b)====
qqnorm(golub[790, gol.fac=="ALL"])
qqline(golub[790, gol.fac=="ALL"])
qqnorm(golub[66, gol.fac=="ALL"])
qqline(golub[66, gol.fac=="ALL"])

#c)====
ds_mean_66<-mean(golub[66,gol.fac=="ALL"])
ds_mean_790<-mean(golub[790,gol.fac=="ALL"])
ds_median_66<-median(golub[66,gol.fac=="ALL"])
ds_median_790<-median(golub[790,gol.fac=="ALL"])

```

