---
title: "Exercise_04
output: html_notebook
---
1. Gene CD33. Use grep to find the index of the important gene CD33
among the list of characters golub.gnames. For each test below for-
mulate the null hypothesis, the p-value and your conclusion.
(a) Test the normality of the ALL and AML expression values.
(b) Test for the equality of variances.
(c) Test for the equality of the means by an appropriate t-test.
(d) Is the experimental effect strong?

```{r}




```

2. Gene ”MYBL2 V-myb avian myeloblastosis viral oncogene homolog-
like 2” has its expression values in row 1788.
(a) Use a boxplot to construct a hypothesis about the experimental
effect.
(b) Test for the equality of means by an appropriate t-test.
```{r}

```

3. HOXA9. Gene ”HOXA9 Homeo box A9” with expression values in row
1391, can cause leukemia (Golub et al., 1999).
(a) Test the normality of the expression values of the ALL patients.
(b) Test for the equality of means by an appropriate t-test.
```{r}

```

4. Zyxin. On NCBI there are various cDNA clones of zyxin.
(a) Find the accession number of cDNA clone with IMAGE:3504464.
(b) Test whether the frequencies of the nucleotides are equal for each
nucleic acid.
(c) Test whether the frequencies of ”X94991.1” can be predicted by
the probabilities of the cDNA sequence ”BC002323.2”.
```{r}

```
5. Gene selection. Select the genes from the golub data with smallest
two-sample t-test values for which the ALL mean is greater than the
AML mean. Report the names of the best ten. Scan the Golub (1999)
article for genes among the ten you found and discuss their biological
function briefly.
```{r}

```
6. Antigenes. Antigenes play an important role in the development of
cancer. Order the antigenes according to their p-values from the Welch
two-sample t-test with respect to gene expression values from the ALL
and AML patients of the Golub et al. (1999) data.
```{r}

```
7. Genetic Model. A certain genetic model predicts that four phenotypes
occur in ration 9:3:3:1. In a certain experiment the offspring is observed
with frequencies 930, 330, 290, 90. Do the data confirm the model?
```{r}

```

8. Comparing two genes. Consider the gene expression values in row 790
and 66 of the Golub et al. (1999) data.
(a) Produce a boxplot for the ALL expression values and comment on
the differences. Are there outliers?
(b) Compute the mean and the median for the ALL gene expression
values for both genes. Do you observed difference between genes?
(c) Compute three measures of spread for the ALL expression values
for both genes. Do you observe difference between genes?
(d) Test by Shapiro-Wilk and Anderson-Darling the normality for the
ALL gene expression values for both genes.
```{r}

```

9. Normality tests for gene expression values of the Golub et al. (1999)
data. Perform the Shapiro-Wilk normality test separately for the ALL
and AML gene expression values. What percentage passed the normal-
ity test separately for the ALL and the AML gene expression values?
What percentage passes both testes?
```{r}

```
10. Two-sample tests on gene expression values of the Golub et al. (1999)
data.
(a) Perform the two-sample Welch t-test and report the names of the
ten genes with the smallest p-values.
(b) Perform the Wilcoxon rank test and report the names of the ten
genes with the smallest p-values.
```{r}

```
11. Biological hypotheses. Suppose that the probability to reject a biolog-
ical hypothesis by the results of a certain experiment is 0.05. Suppose
that the experiment is repeated 1000 times.
(a) How many rejections do you expect.
(b) What is the probability of less than 10 rejections?
(c) What is the probability of more than 5 rejections?
(d) What is the probability that the number of rejections is between
two and eight?
```{r}

```

12. Programming some tests.
(a) Program the two-sample t-test with equal variances and illustrate
it with the expression values of row 1024 the of Golub et al. (1999)
data.
(b) The value of W in the two-sample Wilxoxon test equals the sum
of the ranks of Group 1 minus n(n + 1)/2, where n is the number
of gene expression values in Group 1. Program this and illustrate
it with the expression values of row 1024 of Golub et al. (1999)
data.
(c) The value of W in the two-sample Wilxoxon test equals the num-
ber of values x i > y j , where x i , y j are values from Group 1 and
2, respectively. Program this and illustrate it with the expression
values of row 1024 of Golub et al. (1999) data.