---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
1. Binomial Let X be binomially distributed with n = 60 and p = 0.4.
Compute the following.
(a) P (X = 24), P (X ≤ 24), and P (X ≥ 30).
(b) P (20 ≤ X ≤ 30), P (20 ≤ X).
(c) P (20 ≤ X or X ≥ 40), and P (20 ≤ X and X ≥ 10).
(d) Compute the mean and standard deviation of X.
(e) The quantiles x 0.025 , x 0.5 , and x 0.975 .

```{r}

probability <- rep(0.4, 60) 
plot(probability,
     xlab = "outcomes",
     main = "Probability Distribution") 
#a)====
dbinom(24,60,0.4)
pbinom(24,60,0.4)
1-pbinom(29,60,0.4)
sum(dbinom(30:60,60,0.4))

#b)====
sum(dbinom(20:30,60,0.4))
1-pbinom(19,60,0.4)
#P(X>=20)+P(X>=40)
sum(dbinom(20:60,60,0.4))+sum(dbinom(40:60,60,0.4))
#c)====
#P(X>=20) AND P(X>=10)
sum(dbinom(20:60,60,0.4))
sum(dbinom(10:60,60,0.4))
#Da 20:60 bereits in 10:60 enthalten ist, gilt nur 10:60
#d)====
#Erwartungswert ist SUM(Xxf(x))
#every value has the same probability
sum(probability)
sqrt(60*0.4*0.6)

#c)====
qbinom(c(0.025,0.5,0.975),60,0.4)
```

2. Standard Normal. Compute the following probabilities and quantiles.
(a) P (1.6 < Z < 2.3).
(b) P (Z < 1.64).
(c) P (−1.64 < Z < −1.02).
(d) P (0 < Z < 1.96).
(e) P (−1.96 < Z < 1.96).
(f) The quantiles z 0.025 , z 0.05 , z 0.5 , z 0.95 , and z 0.975 .
```{r}
#Normal
#µ, σ
#dnorm(x, µ, σ) pnorm(x, µ, σ) qnorm (α, µ, σ) rnorm(10, µ, σ)
#a)====
#P (X < 1.4) = pnorm(1.4, 1.9, 0.5) = 0.1586.
#P(X<2.3)-P(X=1.6)
pnorm(2.3,0,1)-pnorm(1.6,0,1)

#b====

pnorm(1.64,0,1)

#c====
pnorm (-1.02,0,1)-pnorm(-1.64,0,1)

#d====

pnorm(1.96,0,1)-0.5

#e====
pnorm(1.96,0,1)-pnorm(-1.96,0,1)

#f====
qnorm(c(0.025, 0.05,0.5,0.95,0.975))
```
Normal. Compute for X distributed as N (10, 2) the following proba-
bilities and quantiles.
a) P (X < 12).
(b) P (X > 8).
(c) P (9 < X < 10, 5).
(d) The quantiles x 0.025 , x 0.5 , and x 0.975 .
```{r}
#a====
#calculate mean first and sd first

pnorm(12,10,2)

#b====
1-pnorm(8,10,2)
#c====
pnorm(10.5, 10, 2)- pnorm(9,10,2)
#d====
qnorm(c(0.025, 0.05,0.5,0.95,0.975),mean =10, sd=2 )

```
T -distribution. Verify the following computations for the T 6 distribu-
tion.
(a) P (T 6 < 1).
(b) P (T 6 > 2).
(c) P (−1 < T 6 < 1).
(d) P (−2 < T 6 < −2).
(e) The quantiles t 0.025 , t 0.5 , and t 0.975 .
```{r}
#T #m=6 #dt(x, m) #pt(x, m) #qt(α, m) #rt(10, m)
#a====
pt(1,6)
#b====
1-pt(2,6)
#c====
pt(1,6)-pt(-1,6)
#d==== Rechtschreibfehler?
pt(2,6)-pt(-2,6)
#e====
qt(c(0.025, 0.05,0.5,0.95,0.975),6 )

```
F distribution. Compute the following probabilities and quantiles for
the F (8,5) distribution.
(a) P (F 8,5 < 3).
(b) P (F 8,5 > 4).
(c) P (1 < F 8,5 < 6).
(d) The quantiles f 0.025 , f 0.5 , and f 0.975 .
```{r}
#m,n df(x, m, n) pf(x, m, n) qf(α, m, n) rf(10, m, n)
#a)====
pf(3, 8,5)
#b)====
1-pf(4,8,5)
#c)====
pf(6,8,5)-pf(1,8,5)

#d)====
qf(c(0.025, 0.05,0.5,0.95,0.975),8,5 )


```
Chi-squared distribution. Compute the following for the chi-squared
distribution with 10 degrees of freedom.
(a) P (χ^2^10 < 3).
(b) P (χ^2^10 > 4).
(c) P (1 < χ^2^10 < 6).
(d) The quantiles g 0.025 , g 0.5 , and g 0.975 .
```{r}
#Chi-squared m dchisq(x, m) pchisq(x, m) qchisq(α, m) rchisq(10, m)
#a)====

pchisq(3,10)
#b)====
1-pchisq(4,10)
#c)====
pchisq(6,10)-pchisq(1,10)
#d)====
qchisq(c(0.025, 0.05,0.5,0.95,0.975),10)

```
MicroRNA. Suppose that for certain microRNA of size 20 the proba-
bility of a purine is binomially distributed with probability 0.7.
(a) What is the probability of 14 purines?
(b) What is the probability of less than or equal to 14 purines?
(c) What is the probability of strictly more than 10 purines?
(d) By what probability is of the number of purines between 10 and
15?
(e) How many purines do you expect? In other words: What is the
mean of the distribution?
(f) What is the standard deviation of the distribution?
```{r}
#a)====
#Use binomial Distribution: pbinom(x,n,p)
#P(X=14)
dbinom(14,20, 0.7)
#b)====
#P(X<= 14)
pbinom(14,20,0.7)

#c)====
#P(X>10)
1-pbinom(10,20,0.7)

#d)====
#P(10<X<15)
pbinom(15,20,0.7)-pbinom(10,20,0.7)

#e)====
#m=N*p
20*0.7

#f)====
#square root of variance S^2: S^2=Np(1-p)
sqrt(20*0.7*(1-0.7))

```
Zyxin. The distribution of the expression values of the ALL patients
on the Zyxin gene are distributed according to N (1.6, 0.4^2 ).
(a) Compute the probability that the expression values are smaller
than 1.2?
(b) What is the probability that the expression values are between 1.2
and 2.0?
(c) What is the probability that the expression values are between 0.8
and 2.4?
(d) Compute the exact values for the quantiles x 0.025 and x 0.975 .
(e) Use rnorm to draw a sample of size 1000 from the population and
compare the sample mean and standard deviation with that of the
population.
```{r}
#Use normal distribution
#a)====
#P(X<1.2)
pnorm(1.2,1.6,0.4)

```
Some computations on Golub et al. (1999) data.
(a) Take µ = 0 and compute the t-values for the ALL gene expression
values. Find the three genes with largest absolute t-values.
(b) Compute per gene the ratio of the variances for the ALL and the
AML patients. How many are between 0.5 and 1.5?
```{r}

```
Extreme value investigation. This (difficult!) question aims to teach
the essence of an extreme value distribution! An interesting extreme
value distribution is given by Pevsner (2003, p.103). Take the maximum
of a sample (with size 1000) from the standard normal distribution and
repeat this 1000 times. So that you sampled 1000 maxima. Next,
subtract from these maxima an and divide by bn, where
```{r}
an <- sqrt(2*log(n)) - 0.5*(log(log(n))+log(4*pi))*(2*log(n))^(-1/2)
bn <- (2*log(n))^(-1/2)


```
Now plot the density from the normalized maxima and add the extreme
value function f (x) from Pevsner his book, and add the density (dnorm)
from the normal distribution. What do you observe?
```{r}

```
