---
title: "R Notebook"
output: html_notebook
---
Z-Test
Example 1. To illustrate the Z-test we shall concentrate on the Gdf5
gene from the Golub et al. (1999) data 2 . The corresponding expression
values are contained in row 2058. A quick search through the NCBI site
makes it likely that this gene is not directly related to leukemia. Hence,
we may hypothesize that the population mean of the ALL expression values
equals zero. Accordingly, we test H 0 : µ = 0 against H 1 : µ 6 = 0. For the sake
of illustration we shall pretend that the standard deviation σ is known to be
equal to 0.25. The z-value (=0.001116211) can be computed as follows.
```{r}
data(golub, package = "multtest")
gol.fac <- factor(golub.cl,levels=0:1, labels= c("ALL","AML"))
sigma <- 0.25; n <- 27; mu0 <- 0
x <- golub[2058,gol.fac=="ALL"]
z.value <- sqrt(n)*(mean(x) - mu0)/sigma

```
The p-value can now be computed as follows. Since it is clearly larger than 0.05, we conclude that the null hypothesis of
mean equal to zero is not rejected (accepted).
```{r}
2*pnorm(-abs(z.value),0,1)
```
Example 2. Using the data from Example 1, the 95% confidence interval
given by Equation 4.1 can be computed as follows. 4
```{r}
mean(x)+qnorm(c(0.025),0,1)*sigma/sqrt(n)
mean(x)+qnorm(c(0.975),0,1)*sigma/sqrt(n)
```
```{r}
library(TeachingDemos)
z.test(x,mu=0,sd=0.25)
```

```{r}
ci.examp(mean.sim =0, sd = 1, n = 25, reps = 100,method = "z", lower.conf=0.025, upper.conf=0.975)
```
 T-Value Test: 
```{r}
x <- golub[2058,gol.fac=="ALL"]; mu0 <- 0; n <- 27
t.value<-sqrt(n)*(mean(x) - mu0)/sd(x)
t.value

#2 · P (T26 ≤ −0.0010) = 2 ∗ pt(−0.0010, 26) = 0.9991 > α,
#To see whether the observed t-value belongs to the 95% confidence inter-
#val, we compute
#(t 0.025,26 , t 0.975,26 ) = (qt(0.025, n − 1), qt(0.975, n − 1)) = (−2.055, 2.055).
```
Since this interval does contain the t-value, we do not reject the hypothesis
that µ equals zero. The left boundary of the 95% confidence interval for the
population mean can be computed, as follows.
```{r}
mean(x)+qt(0.025,26)*sd(x)/sqrt(n)
```
The 95% confidence interval equals (−0.1025, 0.1025). Since it contains zero,
we do not reject the null-hypothesis.
In daily practice it is much more convenient to use the built-in-function
t.test. We illustrate it with the current testing problem.
```{r}
t.test(x,mu=0)

t.test(golub[1042,gol.fac=="ALL"],mu=0, alternative = c("greater"))
```
The large t-value indicates that, relative to its standard error, the mean dif-
fers largely from zero. Accordingly, the p-value is very close to zero, so that
the conclusion is to reject H 0
```{r Two-sample t-test with unequal variances}
#Welch Two-Sample Test
t.test(golub[1042,] ~gol.fac, var.equal=FALSE)
```
The t-value is quite large, indicating that the two means x and y differ largely
from zero relative to the corresponding standard error (denominator in Equa-
tion 4.2). Since the p-value is extremely small, the conclusion is to reject the
null-hypothesis of equal means.
```{r Two sample t-test with equal variances}
t.test(golub[1042,] ~gol.fac, var.equal = TRUE)
```
-> Conclusion is to reject the null Hypothesis of equal population means. P-value is slightly smaller than that of the previous test. 
```{r F-test on equal variances}
var.test(golub[1042,] ~gol.fac)
```
From the p-value 0.4652, the null-hypothesis of equal variances is not re-
jected.
```{r Binomial Test}
binom.test(18, 22, p = 0.7, alternative = c("greater"), conf.level = 0.95)
```
```{r}
library(ape)
zyxinfreq <- table(read.GenBank(c("X94991.1"),as.character=TRUE))
chisq.test(zyxinfreq)
```

```{r}
dat <- matrix(c(5,5,5,5),2,byrow=TRUE)
chisq.test(dat)

dat <- matrix(c(300,500,3000,6000),2,byrow=TRUE)
fisher.test(dat)
```

To test the hypothesis that the ALL gene expression values
of CCND3 Cyclin D3 from Golub et al. (1999) are normally distributed, the
Shapiro-Wilk test can be used as follows.


```{r Normality Test}
shapiro.test(golub[1042, gol.fac=="ALL"])

library(nortest)
ad.test(golub[1042,gol.fac=="ALL"])
```

Since the p-value is greater than 0.05, the conclusion is not to reject the null
hypothesis that CCND3 Cyclin D3 expression values follow from a normal
distribution.

```{r OUtlier Tests}
library(outliers)
grubbs.test(golub[1042, gol.fac=="ALL"])
```

```{r Wilcoxon Rank Test}
wilcox.test(golub[1042,] ~gol.fac)
```
Application on Datasets

```{r}
 data(golub,package="multtest")
 gol.fac <- factor(golub.cl,levels=0:1, labels= c("ALL","AML"))
 sh <- apply(golub[,gol.fac=="ALL"], 1, function(x) shapiro.test(x)$p.value)
 sum(sh > 0.05)/nrow(golub) * 100
```
The Shapiro Test tests for normality. Here 58 % of all ALL data is normaliy distributed -> 42% are not normally distributed! 

In case the gene expression data are non-normally dis-
tributed the t-test may indicate conclusions different from those of the Wilcoxon
test. Differences between these can be investigated by collecting the p-values
from both tests and seeking for the largest differences.
```{r}
data(golub, package = "multtest");
gol.fac <- factor(golub.cl,levels=0:1, labels= c("ALL","AML"))
pt <- apply(golub, 1, function(x) t.test(x ~ gol.fac)$p.value)
pw <- apply(golub, 1, function(x) wilcox.test(x ~ gol.fac)$p.value)
resul <- data.frame(cbind(pw,pt))
resul[pw<0.05 & abs(pt-pw)>0.2,]
```
The p-value is extracted from the output of the t.test function and stored
in the vector pt. The logical operator & is used to select genes for which the
Wilcoxon p-value is smaller than 0.05 and the absolute difference with the
p-value from the t-test is larger than 0.2. Since there are only two such genes
we can draw the reassuring conclusion that the tests give similar results.
