---
title: "Testscript_02"
output: html_notebook
---
Example 1: 
```{r Read from Genbank}
install.packages(c("ape"),repo="http://cran.r-project.org",dep=TRUE)
library(ape)
table(read.GenBank(c("X94991.1"),as.character=TRUE))
pie(table(read.GenBank(c("X94991.1"),as.character = TRUE)))

```

```{r Jitterplot}
data(golub, package = "multtest")
plot(golub[1042,])
gol.fac <- factor(golub.cl,levels=0:1, labels= c("ALL","AML"))
stripchart(golub[1042,] ~ gol.fac, method="jitter", vertical =TRUE)
```

```{r histogram}
hist(golub[1042, gol.fac=="ALL"])
```


```{r sort}
x <- sort(golub[1042, gol.fac=="ALL"], decreasing = FALSE)
x[1:5]
```


```{r Boxplot}
#----Boxplot----
boxplot(golub[1042,] ~gol.fac)
#contructs sequence running from 0.00 to 1.00 with 0.25 steps
pvec <- seq(0,1,0.25)
quantile(golub[1042, gol.fac=="ALL"],pvec)
min(golub[1042, gol.fac=="ALL"])
max(golub[1042, gol.fac=="ALL"])
range(golub[1042, gol.fac=="ALL"])
#to test for outliers: 
#x < x 0.25 − 1.5 · (x 0.75 − x 0.25 ) or x > x 0.75 + 1.5 · (x 0.75 − x 0.25 ).
#extreme outliers factor 3.0 instead of 1.5

```


```{r Quantile Distribution QQ plot}
#----Quantile Quantile Plot----
#Comparison to normal distribution is made, representated by a straight line 
qqnorm(golub[1042, gol.fac=="ALL"])
qqline(golub[1042, gol.fac=="ALL"])
```

Thus the sample mean is simply the average of the n data values. Since it
is the sum of all data values divided by the sample size, a few extreme data
values may largely influence its size. In other words, the mean is not robust
against outliers.

The median is defined as the second quartile or the 50th percentile, and
is denoted by x 0.50 . When the data are symmetrically distributed around the
mean, then the mean and the median are equal. Since extreme data values
do not influence the size of the median, it is very robust against outliers.
Robustness is important in bioinformatics because data are frequently con-
taminated by extreme or otherwise influential data values.
```{r Mean and Median}
mean(golub[1042, gol.fac=="ALL"])
median(golub[1042, gol.fac=="ALL"])

#- as they differ by little, a symmetric data distribution is likely
```

The most important measures of spread are the standard deviation, the in-
terquartile range, and the median absolute deviation. The standard deviation is the square root
of the sample variance, which is defined as the squared differences between the data values and the sample mean. 
The sample standard deviation s is the square root
of the sample variance and may be interpreted as the distance of the data
values to the mean. The variance and the standard deviation are not robust
against outliers.
The interquartile range is defined as the difference between the third and
the first quartile, that is x 0.75 − x 0.25 . More specifically, the value IQR(x)/1.349 is a robust estimator of the standard deviation.
The median absolute deviation (MAD) is defined as
a constant times the median of the absolute deviations of the data from the
median (e.g. Jurečková & Picek, 2006, p. 63). Constant in R is 1.4862. It equals the standard deviation in case the data come from a bell-shaped (normal) distribution. Because the interquartile range and the median absolute deviation are based on quantiles, these are robust against outliers.

```{r}
sd(golub[1042, gol.fac=="ALL"])
IQR(golub[1042, gol.fac=="ALL"]) / 1.349
mad(golub[1042, gol.fac=="ALL"])
```