---
title: "R Notebook"
output: html_notebook
---

Choose(x,y) Function: Probability that out of x, y are true with binomial coefficient

For a binomially distributed variable np is the mean, np(1 − p) the variance, and squareRoot(np(1 − p)) the standard deviation.
#build in function
dbinom(k,n,p)
distribution function which gives the cumulative probabilities:
pbinom(k,n,p)

```{r}
library(TeachingDemos)

vis.binom()

choose(3,1)*0.25^1*0.75^2

for (k in 0:3) print(dbinom(k,3,0.25))

pbinom(2,3,0.25)


```
Example 3. RNA consists of a sequence of nucleotides A, G, U, and C,
where the first two are purines and the last two are pyrimidines. Suppose, for
the purpose of illustration, that the length of a certain micro RNA is 22, that
the probability of a purine equals 0.7, and that the process of placing purines
and pyrimidines is binomially distributed. The event that our microRNA
contains 14 purines can be represented by X = 14. The probability of this
event can be computed by Paper Site 32

By the first line the sequence of integers {1, 2, · · · , 22} is constructed and by
the second the density function is plotted, where the argument h specifies
pins. From Figure 3.1 it can be observed that the largest probabilities oc-
cur near the expectation 15.4. The graph in Figure 3.2 illustrates that the
distribution is an increasing step function, with x on the horizontal axis and
P (X ≤ x) on the vertical.
A random sample of size 1000 from the binomial distribution with n = 22
and p = 0.7 can be drawn by the command rbinom(1000,22,0.7). This
simulates the number of purines in 1000 microRNA’s each with purine prob-
ability equal to 0.7 and length 22.
```{r}
#plotting the binomial distribution and the cumulative properties (density)====
x<-0:22
plot(x, dbinom(x, size=22,prob=.7),type="h")

```

```{r}
library(TeachingDemos)
vis.normal()
vis.t()
pchisq(8,5)

qchisq(0.025, 5, lower.tail=TRUE)
```


```{r}
library(multtest); data(golub)
gol.fac <- factor(golub.cl,levels=0:1, labels= c("ALL","AML"))
x <- golub[1042,gol.fac=="ALL"]
z <- (x-1.90)/0.50
sum(z^2)
#Chi-Square Distribution ====
pchisq(sum(z^2),27, lower.tail=FALSE)
#t-Distribution====
n <- 11
x <- golub[2058, gol.fac=="AML"]
t.value <- sqrt(n)*(mean(x)-0)/sd(x)
t.value
qt(0.025,n-1)

#FDistribution====
var(golub[1042,gol.fac=="ALL"])/var(golub[1042,gol.fac=="AML"])
qf(.025,26,10)

#Plotting density function ====
f<-function(x){dnorm(x,1.9,0.5)}
plot(f,0,4,xlab="x-axis",ylab="density f(x)")

plot(f,0,4,xlab="x-axis",ylab="density f(x)")
x<-seq(0,1.4,0.01)
polygon(c(0,x,1.4), c(0,f(x),0), col="lightblue")

```
For practical computations R has built-in-functions for the binomial, normal,
t, F, χ 2 -distributions, where d stands for density, p for (cumulative) prob-
ability distribution, q for quantiles, and r for drawing random samples.
